
<!--
To compile run:
library(rmarkdown); render( "SimsZvals.Rmd" );knitr::purl("SimsZvals.Rmd")
-->



Simulation using normal mixtures
======================

 **LAST UPDATE AT**

```{r, echo=FALSE}
print(date())
```

Preparations
--------------

We first set global chunk options and load the
neccessary packages.

```{r setup}
library(knitr)
library(reshape2)
library(tidyverse)
library(concaveFDR)
library(ggplot2)
library(grid)
library(locfdr)
library(fdrtool)
library(devtools)

if (!("mixfdr" %in%  (installed.packages()[, 1]))) {
  devtools::install_url("https://cran.r-project.org/src/contrib/Archive/mixfdr/mixfdr_1.0.tar.gz")
}

library(mixfdr)
library(BiocParallel)
library(grid)
library(RColorBrewer)

opts_chunk$set(fig.width=14, fig.height=10, cache=T, 
error=FALSE, cache.lazy=FALSE)
options(digits = 2, scipen = 10)

multicoreParam <- MulticoreParam(workers = round(parallel::detectCores() * 0.75),
                                 catch.errors = FALSE)
multicoreParam
register(multicoreParam)

```

## Set parameters for the simulation

We now set parameters to simulate "z-scores" according to the mixture models
of Strimmer 2008 and create some helper functions. 

```{r setting parameters}

### number of reps
B = 1000 
### number of stats
d = 200
### prop of null
eta0 = 0.8
### sigma
sigma = 2

```
We have a total of `d ` z-values
and we repeat the simulation `B ` times (B=`r B`)

## Simulate z scores

We now  draw the mixture model test statistics.

```{r simulate z scores, eval=FALSE}
##  d z-scores in every column 

  ZMatEasy  <- t(laply(bplapply(seq_len(B), function(i){ 
   get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 5, alt.max = 10)$z 
	}), identity))

#set.seed(777)
#hist(get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 5, alt.max = 10)$z)

save(ZMatEasy, file = "ZMatEasy.rda")


  ZMatHard  <- t(laply(bplapply(seq_len(B), function(i){ 
   get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 2, alt.max = 10)$z 
  }), identity))

#
#hist(get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 2, alt.max = 10)$z)

save(ZMatHard, file = "ZMatHard.rda")

```
Load the presimulated HND scores
```{r load simulated HND scores, eval=TRUE}
load("ZMatEasy.rda")
z <- ZMatEasy[,1]
hist(z)
load("ZMatHard.rda")
z2 <- ZMatHard[,1]
hist(z2)
```

## Helper functions

The following helpfer functions run various FDR estimation algorithms.


```{r helper functions}

#### z-score based FDR methods ####

run.fdrtool = function(z, type=1)
{
  require("fdrtool") 
  if(type == 1) res = fdrtool(z, "normal", plot=FALSE, verbose=FALSE)
  if(type == 2) res = fdrtool(z, "normal", plot=FALSE, verbose=FALSE, cutoff.method="pct0", pct0=0.75)
  if(type == 3) res = fdrtool(z, "normal", plot=FALSE, verbose=FALSE, cutoff.method="locfdr")
  
  if(type == 4){
   source("/home/bklaus/uni/Projects/fdrtool-dev/TESTConcave/log.fdr.R")
   resFDR = log.fdr.fda(z, theo = FALSE, classic = FALSE)
   eta0 =  resFDR$eta0
   sd = resFDR$sd

   res   = fdrtool(resFDR$pval, "pvalue", plot=FALSE, verbose=FALSE)
   }
 
 if(type != 4){	
  eta0 = res$param[,"eta0"]
  sd = res$param[,"sd"]
  }
  fdr = res$lfdr
  Fdr = res$qval
 
  return( list(eta0=eta0, sd=sd, fdr=fdr, Fdr=Fdr) )
}


run.locfdr = function(z)
{
  require("locfdr")
  res = locfdr(z, plot=0)

  eta0 = res$fp0["mlest","p0"]
  sd = res$fp0["mlest","sigma"]
  fdr = res$fdr
  Fdr = NULL

  return( list(eta0=eta0, sd=sd, fdr=fdr, Fdr=Fdr) )
}


run.ConcaveFDR = function(z){

  res = ConcaveFDR(z, "normal",plot=FALSE, verbose =FALSE, cutoff.method = 
  "smoothing")  
  fdr = res$lfdr.log
  Fdr = res$qval.log
  
  return( list(eta0=res$param[3], sd=res$param[5], fdr=fdr, Fdr=Fdr) )
}



```

## Run FDR estimation algorithms

simulation helper functions
```{r simulation helper functions}

simFDREasy = function(method, ...)
{
  
  return(ldply(bplapply(seq_len(B), function(i){ 
    z <- ZMatEasy[,i]
    tmp <-  get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 5, alt.max = 10)
     true.Fdr = tmp$Fdr.func(z)
     true.fdr = tmp$fdr.func(z)
     out = method(z, ...)
     # out = method(z)
    res = numeric(6)
    names(res) = c("eta0", "sd", "fdr.MSE", "Fdr.MSE", "fdr.MSE.subset", "Fdr.MSE.subset")
    
    res[ 1] = out$eta0
    res[ 2] = out$sd
    res[ 3] = mean( (true.fdr-out$fdr)^2 )
    res[ 4] = mean( (true.Fdr-out$Fdr)^2 )

    idx = (abs(z)>2)


    res[ 5] = mean( ((true.fdr-out$fdr)[idx])^2 )
    res[ 6] = mean( ((true.Fdr-out$Fdr)[idx])^2 )
    
    return(res)
     }),identity))
   
}

simFDRHard = function(method, ...)
{
  
  return(ldply(bplapply(seq_len(B), function(i){ 
    z <- ZMatHard[,i]
    tmp <-  get.random.zscore(d, sigma = sigma, eta0 = eta0, alt.min = 2, alt.max = 10)
     true.Fdr = tmp$Fdr.func(z)
     true.fdr = tmp$fdr.func(z)
     out = method(z, ...)
     # out = method(z)
    res = numeric(6)
    names(res) = c("eta0", "sd", "fdr.MSE", "Fdr.MSE", "fdr.MSE.subset", "Fdr.MSE.subset")
    
    res[ 1] = out$eta0
    res[ 2] = out$sd
    res[ 3] = mean( (true.fdr-out$fdr)^2 )
    res[ 4] = mean( (true.Fdr-out$Fdr)^2 )

    idx = (abs(z)>2)

    res[ 5] = mean( ((true.fdr-out$fdr)[idx])^2 )
    res[ 6] = mean( ((true.Fdr-out$Fdr)[idx])^2 )
    
    return(res)
     }),identity))
   
}
```

We can now run the estimation algorithms and save the results.

```{r run simulations, eval = FALSE}
EasyResults <- ldply(list(ConcaveFDR = simFDREasy(run.ConcaveFDR),
              locfdr = simFDREasy(run.locfdr),
              fdrtool = simFDREasy(run.fdrtool)), identity, .id = "Method")
               

HardResults <- ldply(list(ConcaveFDR = simFDRHard(run.ConcaveFDR),
              locfdr = simFDRHard(run.locfdr),
              fdrtool = simFDRHard(run.fdrtool)), identity, .id = "Method")

#test <- simFDREasy(run.ConcaveFDR)
#test2 <- simFDRHard(run.ConcaveFDR)

save(EasyResults, HardResults, file = "ZModelResults.RData")

```

```{r load simulations results}
load("ZModelResults.RData")
#display.brewer.pal(6, name ="Set3")


```
## Produce plots

Here we produce some plots to evaluate the estimation  results graphically.
We first define some helper functions for the plots

```{r helper functions for plots}
	vplayout <- function(x, y){
	viewport(layout.pos.row = x, layout.pos.col = y)
	}

#locfdrMLE fdrtoolEmp resConcave 
# "#D95F02"  "#7570B3"  "#E6AB02" 

 #pal <- brewer.pal(3, name ="Dark2")
  pal <- c("#E6AB02" , "#D95F02" , "#7570B3" )
names(pal) <- c("ConcaveFDR", "locfdr" ,"fdrtool")
```


### fdr accuracy plots
```{r plots - fdr accuracy}
(qplot(Method, fdr.MSE, data = EasyResults, color = Method, main = "MSE Easy Model") 
    + geom_boxplot() 
    + ylim(0,.1) + scale_color_manual(values = c(pal)))

(qplot(Method, fdr.MSE, data = HardResults, color = Method, main = "MSE Hard Model") 
    + geom_boxplot() 
    + ylim(0,.1) + scale_color_manual(values = c(pal)))
```


### fdr accuracy plots subset with fdr < 0.2
```{r plots - fdr accuracy subset}

(qplot(Method, fdr.MSE.subset, data = EasyResults, color = Method, main = "MSE Subset Easy Model") 
    + geom_boxplot() 
    + ylim(0,.1) + scale_color_manual(values = c(pal)))

(qplot(Method, fdr.MSE.subset, data = HardResults, color = Method, main = "MSE Subset Hard Model") 
    + geom_boxplot() 
    + ylim(0,.1) + scale_color_manual(values = c(pal)))
```


### Null model parameters accuracy plots

## eta0

```{r plots - eta0 accuracy}
(qplot(Method, eta0, data = EasyResults, color = Method, main = "eta0 Easy Model")
    + geom_boxplot() + ylim(0.6,1)
  + scale_color_manual(values = c(pal)) + 
   geom_hline(aes(yintercept = get("eta0", envir=  .GlobalEnv )), 
              color ="coral3")
 )

(qplot(Method, eta0, data = HardResults, color = Method, main = "eta0 Hard Model")
    + geom_boxplot() + ylim(0.6,1)
  + scale_color_manual(values = c(pal)) + 
   geom_hline(aes(yintercept = get("eta0", envir=  .GlobalEnv )), 
              color ="coral3")
 )


```


## sigma


```{r plots - sigma accuracy}
# ## get accuracy for all methods
(qplot(Method, sd, data = EasyResults, color = Method, main = "sigma Easy Model")
    + geom_boxplot() + ylim(1,3.5)
  + scale_color_manual(values = c(pal)) + 
   geom_hline(aes(yintercept = get("sigma", envir=  .GlobalEnv )), 
              color ="coral3")
 )

(qplot(Method, sd, data = HardResults, color = Method, main = "sigma Hard Model")
    + geom_boxplot() + ylim(1,3.5)
  + scale_color_manual(values = c(pal)) + 
   geom_hline(aes(yintercept = get("sigma", envir=  .GlobalEnv )), 
              color ="coral3")
 )

```


 **SESSION INFO**
 

```{r SessionInfo, echo=FALSE}

sessionInfo()

```

